Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-05-01 manuscript/manuscript.tex
msupdates

@@

  \caption{$G(t)$ trajectories from background selection simulations both
  including fixed sites (solid lines) and not including fixed sites (dashed
lines). The dashed lines are identical to those in Supplementary Materials
Figure \ref{suppfig:supp-bgs-g}. As expected, including fixed sites shrinks the
$G(t)$ towards zero, and in some cases, they can become negative.}

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.6\textwidth]{figures/figure-3-hists-b-without-fixations.pdf}

  \caption{}

>>>>

  \caption{$G(t)$ trajectories from background selection simulations both
  including fixed sites (solid lines) and not including fixed sites (dashed
lines). The dashed lines are identical to those in Supplementary Materials
Figure \ref{suppfig:supp-bgs-g}. }

We noticed in some simulations, when temporal covariances are calculated over
longer time periods (i.e. $>50$ generations), that neutral simulations appear
to slowly trend towards negative $G(t)$ (Supplementary Materials Figure
\ref{suppfig:supp-bgs-g-fix}). In investigating this bias, we found that the
effect went away when we marked fixed or lost sites as missing, and calculated
the temporal covariance; furthermore, this affected the $G(t)$ trajectories
under selection too (Supplementary Material Figure
\ref{suppfig:supp-bgs-g-fix}).

Looking a neutral simulations results, we found that spuriously significant
$G(t)$ trajectories (e.g. different from zero under the case of neutrality) can
be created by conditioning on sojourn time. Our intuition, which will need to
be confirmed by future work and theory, is that by conditioning on a quick
sojourn time to fixation (or loss) from an initial intermediate frequency,
leads to trajectories that, purely by chance, have many upward (or downward)
allele frequency changes, which leads to a positive sum of pairwise temporal
covariances (the numerator of $G(t)$). Similarly, conditioning on long sojourn
times implies that frequency changes alternate back and forth quite a bit, such
the sum of pairwise temporal covariances could be slightly negative.

When all pairwise covariances are calculated and
summed to calculate the numerator of $G(t)$, this leads to this subset of
neutral trajectories to have spuriously significant and positive $G(t)$
trajectories. We briefly confirmed that conditioning on fixation leads to a
spurious positive numerator of $G(t)$. Looking at one neutral simulation, we
conditioned on alleles fixing during the 105 generations of which temporal
covariances are calculated. Per locus, we calculated the sum of all pairwise
products between allele frequency changes (analogous to the numerator of
$G(t)$). Whether we condition on fixation, or a large absolute allele frequency
change between the initial generation and the last generation, we find that
quickly fixing sites and sites with large allele frequency shifts have a
significant positive sum of covariances as expected, despite being neutral
trajectories. Thus, even under neutrality, one must mind how fixations are
handled.

Since our simulation allele
frequencies are population, rather than sample frequencies, when an allele is
fixed or lost is known with certainty.

@@

@@

\begin{figure}[!ht]
  \centering
  \includegraphics{figures/figure-3-hists-b-without-fixations.pdf}

  \caption{A version of Figure \ref{fig:figure-3} (A) and (B) excluding fixed
    and lost sites. The same qualitative pattern holds as the original figure,
    which did not exclude fixed and lost sites: there is an enrichment of
    positive temporal covariances between near timepoints ($k=2$) in the
  \textcite{Barghi2019-qy} study, and an excess of negative temporal
covariances at more distant timepoints ($k=2$). }
